Author disambiguation of large-scale semi-labeled collections of scientific articles

### 2015-08-20 paper.tex
save

\textit{Input features.} In most cases, supervised learning algorithms assume
the input space ${\cal X}$ to be numeric (e.g., $\mathbb{R}^p$), making them
not directly applicable to structured input spaces such as ${\cal S} \times
{\cal S}$. Following previous works \glnote{add references}, pairs of
signatures $(s_1, s_2)$ are first transformed to vectors $v \in \mathbb{R}^p$
by building so-called similarity profiles
\citep{treeratpituk2009disambiguating}, and on which supervised learning is
then carried on. In this work, we design and evaluate 15 standard input
features based on the comparison of signature fields, as reported in the first
half of Table~\ref{table:features}. As an illustrative example, the
\textit{Full name} feature corresponds to the similarity between the (full)
author name fields of the two signatures, as measured using as combination
operator the cosine similarity between their respective $(n,m)$-TF-IDF vector
representations\footnote{$(n,m)$ denotes that the TF-IDF vectors are computed
from character $n$, $n+1$, ..., $m$-grams. When not specified, TF-IDF vector
are otherwise computed from words.}. Similarly, the \textit{Year difference}
feature measures the absolute difference between the publication date of the
articles to which the two signatures respectively belong.

Observing that author names from different cultures or origins are likely to be
disambiguated using different strategies (e.g., pairs of signatures with French
author names versus pairs of signatures with Chinese author names), and as
corroborated in \citep{treeratpituk2012name} or \citep{chin2014effective}, we
propose to complement our feature set with 7 additional new features, each
evaluating the degree of belonging of both signatures to several ethnic groups,
as reported in the second half of Table~\ref{table:features}. More
specifically, using census data extracted from \citep{rugglesintegrated}, we
build a logistic regression classifier for mapping the $(1,5)$-TF-IDF representation of
an author name to one of the 7 ethnic groups reported in that dataset. Given a
pair of signatures $(s_1, s_2)$, the proposed ethnicity features are each
computed as the estimated probability of $s_1$ to belong to the respective
ethnic group, multiplied by the estimated probability of $s_2$ to belong to the
same group. In doing so, the expectation is for the linkage function to become
sensitive to the actual origin of the authors depending on the values of these
features.

@@

% distance learning as a supervised problem
% describe all features
% balancing difficult and easy cases
% sampling pairs from block only

>>>>

\textit{Input features.} In most cases, supervised learning algorithms assume
the input space ${\cal X}$ to be numeric (e.g., $\mathbb{R}^p$), making them
not directly applicable to structured input spaces such as ${\cal S} \times
{\cal S}$. Following previous works \glnote{add references}, pairs of
signatures $(s_1, s_2)$ are first transformed to vectors $v \in \mathbb{R}^p$
by building so-called similarity profiles
\citep{treeratpituk2009disambiguating}, and on which supervised learning is
then carried on. In this work, we design and evaluate 15 standard input
features based on the comparison of signature fields, as reported in the first
half of Table~\ref{table:features}. As an illustrative example, the
\textit{Full name} feature corresponds to the similarity between the (full)
author name fields of the two signatures, as measured using as combination
operator the cosine similarity between their respective $(n,m)$-TF-IDF vector
representations\footnote{$(n,m)$ denotes that the TF-IDF vectors are computed
from character $n$, $n+1$, ..., $m$-grams. When not specified, TF-IDF vectors
are otherwise computed from words.}. Similarly, the \textit{Year difference}
feature measures the absolute difference between the publication date of the
articles to which the two signatures respectively belong.

Observing that author names from different cultures or origins are likely to be
disambiguated using different strategies (e.g., pairs of signatures with French
author names versus pairs of signatures with Chinese author names), and as
corroborated in \citep{treeratpituk2012name} or \citep{chin2014effective}, we
propose to complement our feature set with 7 additional new features, each
evaluating the degree of belonging of both signatures to several ethnic groups,
as reported in the second half of Table~\ref{table:features}. More
specifically, using census data extracted from \citep{rugglesintegrated}, we
build a logistic regression classifier for mapping the $(1,5)$-TF-IDF
representation of an author name to one of the 7 ethnic groups reported in that
dataset. Given a pair of signatures $(s_1, s_2)$, the proposed ethnicity
features are each computed as the estimated probability of $s_1$ to belong to
the corresponding ethnic group, multiplied by the estimated probability of
$s_2$ to belong to the same group. In doing so, the expectation is for the
linkage function to become sensitive to the actual origin of the authors
depending on the values of these features. Indirectly, let us also note that these
features hold discriminative power since, if author names are predicted
to belong to different ethnic groups, then they are also likely to correspond to
distinct actual persons.

@@

% balancing difficult and easy cases
% sampling pairs from block only
